```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(corrplot)
library(leaps)
library(car)
library(Metrics)
library(reshape2)
library(ggpubr)
library(moments)
library(dplyr)
```

```{r}
data <- read.csv("Life Expectancy Data.csv")
head(data)
```

```{r}
missing_counts <- data.frame(feature = factor(names(data)),
                    counts=sapply(data, function(x) sum(is.na(x))))


ggplot(missing_counts,
       aes(x=counts, y=reorder(feature, -counts), fill=counts)) +
                                  geom_bar(stat="identity") +
                                  ggtitle("Missing Value Counts") +
                                  xlab("Feature") + ylab("Missing count") +
                                  theme(axis.text.x=element_text(angle=20, hjust=1))+
                                  theme(text = element_text(size = 12))+
                                  scale_fill_continuous()

```

```{r}
plot(Life.expectancy~Schooling,data=data, main = "Schooling vs. Life Expectancy")
```

```{r}
hist(data$Life.expectancy, 
     xlab = "Life Expectancy",
     main = "Distribution of Life Expectancy")
```
 
##Filling In Missing Values using imputation

```{r}
library(missForest)
```

Need to Drop Country And Year First (Can add them back later)
```{r}
columns_to_exclude <- c("Year", "Country")

# Select all columns except the ones to exclude
le <- data %>%
  select(-one_of(columns_to_exclude))
```

Change necessary variables into factors
```{r}
le$Status <- as.factor(le$Status)
```

The missForest package is specifically designed for imputing missing data in datasets with mixed types of variables (numeric and categorical). It uses a random forest-based imputation method.
```{r}
imputed_data <- missForest(le)

imputed_data <- imputed_data$ximp
```


##Split Data into Training and Testing
```{r}
set.seed(123) 

train_indices <- sample(seq_len(nrow(imputed_data)), size = 0.7 * nrow(imputed_data))

data_train <- imputed_data[train_indices, ]

data_test <- imputed_data[-train_indices, ]
```

##Gabby EDA
Explore Relationship Between Adult Mortality and Life Expectancy (Filled by Status)

Explore Relationship Between Infant Deaths and Life Expectancy (Filled by Status)

Explore Relationship Between Schooling and Life Expectancy (Filled by Status)

Explore Relationship Between BMI and Life Expectancy (Filled by Status)

Explore Relationship Between GDP and Life Expectancy (Filled by Status)

##Emily EDA

##Jessica EDA